Faster generic IND-CCA2 secure KEM using ``encrypt-then-MAC''

### 2024-09-17 etm-transform/manuscript/main.tex
define algorithms as short commands
Reformatted figures and tables

%%%% 1. DOCUMENTCLASS %%%%
%
% TODO: Algo 9, 10, 11 should be one algorithms
%   similarly agorithm 13-21, 22-24, 25-30
% TODO: the implementation should be called ML-KEM+
% TODO: Table 2 should be called ML-KEM+, as well
% TODO: Table 3 should be broken up: pk/sk/keygen measurements should be one table, ct/encap/decap measurements should be one table for clearer comparison "
% TODO: Table 4-6 need double lines to separate security level
% TODO: need more thorough references
% TODO: need better intro and conclusion (but later)
% \documentclass[journal=tches,submission,notanonymous]{iacrtrans}
\documentclass[floatrow,journal=tches,submission]{iacrtrans}
%%%% NOTES:
% - Change "journal=tosc" to "journal=tches" if needed
% - Change "submission" to "final" for final version
% - Add "notanonymous" to reveal authors
% - Add "spthm" for LNCS-like theorems

%%%% 2. PACKAGES %%%%
\usepackage{algorithm}
\usepackage{algpseudocode}
% \usepackage[demo]{graphicx}

@@

\begin{figure}[H]
    \begin{floatrow}
        \ffigbox{
            \begin{algorithm}[H]
                \caption{
                    The $\texttt{OW-ATK}$ game
                }\label{alg:pke-ow-atk-game}
                \begin{algorithmic}[1]
                    \State $(\pk, \sk) \leftsample \keygen(1^\lambda)$
                    \State $m^\ast \leftsample \mathcal{M}$
                    \State $c^\ast \leftsample \encrypt(\pk, m^\ast)$
                    \State $\hat{m} \leftsample \mathcal{A}^{\mathcal{O}_\texttt{ATK}}(1^\lambda, \pk, c^\ast)$
                    \State \Return $\llbrack m^\ast = \hat{m} \rrbrack$
                \end{algorithmic}
            \end{algorithm}
        }
        {\caption{The \texttt{OW-ATK} game}}
        \ffigbox{
            \begin{algorithm}[H]
                \caption{$\pco(m \in \mathcal{M}, c \in \mathcal{C})$}\label{alg:pco}
                \begin{algorithmic}[1]
                    \State \Return $\llbrack \decrypt(\sk, c) = m \rrbrack$
                \end{algorithmic}
            \end{algorithm}
        }
        {\caption{Plaintext-checking oracle}}
    \end{floatrow}
    \label{fig:pco-and-cvo}
\end{figure}

In the \texttt{OW-ATK} game (see figure \ref{alg:pke-ow-atk-game}), an adversary's goal is to recover the decryption of a randomly generated ciphertext. A challenger randomly samples a keypair and a challenge plaintext $m^\ast$, encrypts the challenge plaintext $c^\ast \leftsample \encrypt(\pk, m^\ast)$, then gives $\pk$ and $c^\ast$ to the adversary $A$. The adversary $A$, with access to some oracle $\mathcal{O}_\texttt{ATK}$, outputs a guess decryption $\hat{m}$. $A$ wins the game if its guess $\hat{m}$ is equal to the challenge plaintext $m^\ast$. The \textit{advantage} $\adv_\texttt{OW-ATK}$ of an adversary in this game is the probability that it wins the game:

@@

The capabilities of the oracle $\mathcal{O}_\texttt{ATK}$ depends on the choice of security goal $\texttt{ATK}$. Particularly relevant to our result is security against plaintext-checking attack (PCA), for which the adversary has access to a plaintext-checking oracle (PCO) (see figure \ref{alg:pco}). A PCO takes as input a plaintext-ciphertext pair $(m, c)$ and returns \texttt{True} if $m$ is the decryption of $c$ or \texttt{False} otherwise.

@@

\begin{figure}[H]
    \begin{floatrow}
        \ffigbox{
            \begin{algorithm}[H]
                \caption{
                    \texttt{IND-CCA} game for \kem
                }\label{alg:kem-ind-cca-game}

                \begin{algorithmic}[1]
                    \State $(\pk, \sk) \leftsample \keygen(1^\lambda)$
                    \State $(c^\ast, K_0) \leftsample \encap(\pk)$
                    \State $K_1 \leftsample \mathcal{K}$
                    \State $b \leftsample \{0, 1\}$
                    \State $\hat{b} \leftsample A^{\mathcal{O}_\decap}(
                        1^\lambda, \pk, c^\ast, K_b
                    )$
                    \State \Return $\llbrack \hat{b} = b \rrbrack$
                \end{algorithmic}
            \end{algorithm}
        }{\caption{The \texttt{KEM-IND-CCA2} game}}
        \ffigbox{
            \begin{algorithm}[H]
            \caption{
                $\mathcal{O}_\decap(c)$
            }
            \begin{algorithmic}[1]
                \State \Return $\decap(\sk, c)$
            \end{algorithmic}
            \end{algorithm}
        }{\caption{Decapsulation oracle}}
    \end{floatrow}
\end{figure}

@@

\begin{figure}[H]
    \begin{floatrow}
        \ffigbox[250px]{
            \begin{algorithm}[H]
                \caption{The EUF-CMA game}\label{alg:mac-euf-cma-game}

                \begin{algorithmic}[1]
                    \State $k^\ast \leftsample \mathcal{K}$
                    \State $(\hat{m}, \hat{t}) \leftsample \mathcal{A}^{\mathcal{O}_\sign}()$
                    \State \Return $
                        \llbrack \verify(k^\ast, \hat{m}, \hat{t}) 
                        \land (\hat{m}, \hat{t}) \not\in \mathcal{O}_\sign
                        \rrbrack
                    $
                \end{algorithmic}
            \end{algorithm}
        }{\caption{The \texttt{EUF-CMA} game}}
    \end{floatrow}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \begin{minipage}{0.40\textwidth}
        \begin{algorithm}[H]
            \caption{$\keygen()$}\label{alg:fo-keygen}
            \begin{algorithmic}[1]
                \State $(\pk, \sk^\prime) \leftsample \keygen_\pke()$
                \State $z \leftsample \mathcal{M}$
                \State $\sk \leftarrow (\sk^\prime, z)$
                    \Comment{$U^{\not\bot}, U^{\not\bot}_m$}
                \State $\sk \leftarrow \sk^\prime$
                    \Comment{$U^{\bot}, U^{\bot}_m$}
                \State \Return $(\pk, \sk)$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \hspace{0.05\textwidth}
    \begin{minipage}{0.45\textwidth}
        \begin{algorithm}[H]
            \caption{$\encap(\pk)$}\label{alg:fo-encap}
            \begin{algorithmic}[1]
                \State $m \leftsample \mathcal{M}$
                \State $r \leftarrow G(m)$
                \State $c \leftarrow \encrypt(\pk, m, r)$
                \State $K \leftarrow H(m, c)$
                    \Comment{$U^\bot, U^{\not\bot}$}
                \State $K \leftarrow H(m)$
                    \Comment{$U_m^\bot, U_m^{\not\bot}$}
                \State \Return $(c, K)$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \vspace{0.5cm}
    \begin{minipage}{\textwidth}
        \begin{algorithm}[H]
            \caption{$\decap(\sk=(\sk^\prime, z), c)$}\label{alg:fo-decap}
            \begin{algorithmic}[1]
                \State $\hat{m} \leftarrow \decrypt(\sk^\prime, c)$
                \State $\hat{r} \leftarrow G(\hat{m})$
                \State $\hat{c} \leftarrow \encrypt(\pk, \hat{m}, \hat{r})$
                \If{$\hat{c} = c$}
                    \State $K \leftarrow H(\hat{m})$
                        \Comment{$U_m^\bot, U_m^{\not\bot}$}
                    \State $K \leftarrow H(\hat{m}, c)$
                        \Comment{$U^\bot, U^{\not\bot}$}
                \Else
                    \State $K \leftarrow H(z, c)$
                        \Comment{$U^{\not\bot}, U^{\not\bot}_m$}
                    \State $K \leftarrow \bot$
                        \Comment{$U^{\bot}, U^{\bot}_m$}
                \EndIf 
                \State \Return $K$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \caption{Summary of the modular Fujisaki-Okamoto transformation variations}\label{fig:fo-routines}
\end{figure}

@@

\begin{figure}[H]
    \center
    \begin{algorithm}[H]
        \caption{$\keygen_\etm$}\label{alg:etm-keygen}
        \begin{algorithmic}[1]
            \State $(\pk, \sk^\prime) \leftsample \keygen(1^\lambda)$
            \State $z \leftsample \mathcal{M}$
            \State $\sk \leftarrow (\sk^\prime, z)$
            \State \Return $(\pk, \sk)$
        \end{algorithmic}
    \end{algorithm}
    \begin{algorithm}[H]
        \caption{$\encap(\pk)$}\label{alg:etm-encap}
        \begin{algorithmic}[1]
            \State $m \leftsample \mathcal{M}$
            \State $k \leftarrow G(m)$
            \State $c^\prime \leftsample \encrypt(\pk, m)$
            \State $t \leftarrow \mac(k, c^\prime)$
            \State $K \leftarrow H(m, c^\prime)$
            \State $c \leftarrow (c^\prime, t)$
            \State \Return $(c, K)$
        \end{algorithmic}
    \end{algorithm}
    \begin{algorithm}[H]
        \caption{$\decap(\sk, c)$}\label{alg:etm-decap}
        \begin{algorithmic}[1]
            \State $(c^\prime, t) \leftarrow c$
            \State $(\sk^\prime, z) \leftarrow \sk$
            \State $\hat{m} \leftarrow \decrypt(\sk^\prime, c^\prime)$
            \State $\hat{k} \leftarrow G(\hat{m})$
            \If{$\mac(\hat{k}, c^\prime) \neq t$}
                \State $K \leftarrow H(z, c^\prime)$
            \Else
                \State $K \leftarrow H(\hat{m}, c^\prime)$
            \EndIf
            \State \Return $K$
        \end{algorithmic}
    \end{algorithm}
    \caption{$\kem_\etm$ routines}\label{fig:etm-routines} 
\end{figure}

@@

\begin{figure}[H]
    \begin{floatrow}
        \ffigbox[300px]{
            \begin{algorithm}[H]
                \caption{$\pco(m, c)$}\label{alg:pco-using-decap-oracle}
                \begin{algorithmic}[1]
                    \State $k \leftarrow G(m)$
                    \State $t \leftarrow \mac(k, c)$
                    \State \Return $\llbrack \mathcal{O}^\decap((c, t)) = H(m, c)\rrbrack$
                \end{algorithmic}
            \end{algorithm}
        }{
            \caption{Every decapsulation oracle can be converted into a plaintext-checking oracle}
            \label{fig:pco-using-decap-oracle}
        }
    \end{floatrow}
\end{figure}

@@

\subsection{Proof of theorem \ref{thm:ow-pca-implies-kem-ind-cca2}}
% TODO: write some intuition first?
\begin{proof}
    We will prove theorem \ref{thm:ow-pca-implies-kem-ind-cca2} using a sequence of games.

    \begin{figure}[H]
        \begin{floatrow}
            \ffigbox{
                \begin{algorithm}[H]
                    \caption{IND-CCA2 game for KEM}\label{alg:game-0}
                    \begin{algorithmic}[1]
                        \State $(\pk, \sk) \leftsample \keygen_\etm()$
                        \State $m^\ast \leftsample \mathcal{M}$
                        \State $c^\prime \leftsample \encrypt(\pk, m^\ast)$
                        \State $k^\ast \leftarrow G(m^\ast)$
                        \State $t \leftarrow \mac(k, c^\prime)$
                        \State $c^\ast \leftarrow (c^\prime, t)$
                        \State $K_0 \leftarrow H(m^\ast, c^\prime)$
                        \State $K_1 \leftsample \mathcal{K}_\kem$
                        \State $b \leftsample \{0,1\}$
                        \State $\hat{b} \leftarrow A^{\mathcal{O}^\decap}(\pk, c^\ast, K_b)$
                        \State \Return $\llbrack \hat{b} = b \rrbrack$
                    \end{algorithmic}
                \end{algorithm}
            }{}
            \ffigbox{
                \begin{algorithm}[H]
                    \caption{$\mathcal{O}^\decap(c)$}\label{alg:true-decap-oracle}
                    \begin{algorithmic}[1]
                        \State $(c^\prime, t) \leftarrow c$
                        \State $\hat{m} = \decrypt(\sk^\prime, c^\prime)$
                        \State $\hat{k} \leftarrow G(\hat{m})$
                        \If{$\mac(\hat{k}, c^\prime) = t$}
                            \State $K \leftarrow H(\hat{m}, c^\prime)$
                        \Else 
                            \State $K \leftarrow H(z, c^\prime)$
                        \EndIf 
                        \State \Return $K$
                    \end{algorithmic}
                \end{algorithm}
            }{}
        \end{floatrow}
        \begin{floatrow}
            \ffigbox{
                \begin{algorithm}[H]
                    \caption{$\mathcal{O}^G(m)$}\label{alg:hash-oracle-g}
                    \begin{algorithmic}[1]
                        \If{$\exists (\tilde{m}, \tilde{k}) \in \mathcal{L}^G : \tilde{m} = m$}
                            \State \Return $\tilde{k}$
                        \EndIf
                        \State $k \leftsample \mathcal{K}_\mac$
                        \State $\mathcal{L}^G \leftarrow \mathcal{L}^G \cup \{(m, k)\}$
                        \State \Return $k$
                    \end{algorithmic}
                \end{algorithm}
            }{}
            \ffigbox{
                \begin{algorithm}[H]
                    \caption{$\mathcal{O}^H(m, c)$}\label{alg:hash-oracle-h}
                    \begin{algorithmic}[1]
                        \If{$\exists (\tilde{m}, \tilde{c}, \tilde{K}) \in \mathcal{L}^H : \tilde{m} = m \land \tilde{c} = c$}
                            \State \Return $\tilde{K}$
                        \EndIf
                        \State $K \leftsample \mathcal{K}_\kem$
                        \State $\mathcal{L}^H \leftarrow \mathcal{L}^H \cup \{(m, c, K)\}$
                        \State \Return $K$
                    \end{algorithmic}
                \end{algorithm}
            }{}
        \end{floatrow}
    \end{figure}

@@

    \begin{figure}[H]
        \begin{floatrow}
            \ffigbox[270px]{
                \begin{algorithm}[H]
                    \caption{$\mathcal{O}^\decap_1(c)$}\label{alg:sim-decap-oracle}
                    \begin{algorithmic}[1]
                        \State $(c^\prime, t) \leftarrow c$
                        \If{$\exists (\tilde{m}, \tilde{k}) \in \mathcal{L}^G : 
                            \tilde{m} = \decrypt(\sk^\prime, c^\prime)
                            \land \mac(\tilde{k}, c^\prime) = t
                        $}
                            \State $K \leftarrow H(\tilde{m}, c^\prime)$
                        \Else
                            \State $K \leftarrow H(z, c^\prime)$
                        \EndIf
                        \State \Return $K$
                    \end{algorithmic}
                \end{algorithm}
            }{
                \caption{Simulated decapsulation oracle}
            }
        \end{floatrow}
    \end{figure}

@@

A detailed description of $\texttt{K-PKE}$'s routines can be found in FIPS 203 (TODO: citation). The ``encrypt-then-MAC'' routines are listed in algorithms \ref{alg:ml-kem-etm-keygen}, \ref{alg:ml-kem-etm-encap}, and \ref{alg:ml-kem-etm-decap}.

\begin{algorithm}[H]
    \caption{\texttt{ML-KEM-ETM.KeyGen()}}\label{alg:ml-kem-etm-keygen}
    \begin{algorithmic}[1]
        \State $z \leftsample \{0,1\}^{256}$
        \State $(\pk, \sk^\prime) \leftsample \texttt{K-PKE}.\keygen()$
        \State $h \leftarrow H(\pk)$
            \Comment{$H$ is \texttt{SHA3-256}}
        \State $\sk \leftarrow (sk^\prime \Vert \pk \Vert h \Vert z)$
        \State \Return $(\pk, \sk)$
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}[H]
    \caption{\texttt{ML-KEM-ETM.Encap}(\pk)}\label{alg:ml-kem-etm-encap}
    \begin{algorithmic}[1]
        \Require Public key $\pk$
        \State $m \leftsample \{0,1\}^{256}$
        \State $(\overline{K}, r, k) \leftarrow \texttt{XOF}(m \Vert H(\pk))$
            \Comment{\texttt{XOF} is \texttt{Shake256}, outputting 768 bits}
        \State $c^\prime \leftarrow \texttt{K-PKE}.\encrypt(\pk, m, r)$
        \State $t \leftarrow \mac(k, c^\prime)$
        \State $K \leftarrow \texttt{KDF}(\overline{K} \Vert t)$
            \Comment{\texttt{KDF} is \texttt{Shake256}, outputting 256 bits}
        \State $c \leftarrow (c^\prime, c)$
        \State \Return $(c, K)$
    \end{algorithmic}
\end{algorithm}

>>>>

%%%% 1. DOCUMENTCLASS %%%%
%
% TODO: Algo 9, 10, 11 should be one algorithms
%   similarly agorithm 13-21, 22-24, 25-30
% TODO: the implementation should be called ML-KEM+
% TODO: Table 2 should be called ML-KEM+, as well
% TODO: Table 3 should be broken up: pk/sk/keygen measurements should be one table, ct/encap/decap measurements should be one table for clearer comparison "
% TODO: Table 4-6 need double lines to separate security level
% TODO: need more thorough references
% TODO: need better intro and conclusion (but later)
% \documentclass[journal=tches,submission,notanonymous]{iacrtrans}
\documentclass[journal=tches,submission]{iacrtrans}
%%%% NOTES:
% - Change "journal=tosc" to "journal=tches" if needed
% - Change "submission" to "final" for final version
% - Add "notanonymous" to reveal authors
% - Add "spthm" for LNCS-like theorems

%%%% 2. PACKAGES %%%%
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{caption}
% \usepackage[demo]{graphicx}

@@

\def\algowatkgame{
    \begin{algorithm}[H]
        \caption*{The $\texttt{OW-ATK}$ game}
        \begin{algorithmic}[1]
            \State $(\pk, \sk) \leftsample \keygen(1^\lambda)$
            \State $m^\ast \leftsample \mathcal{M}$
            \State $c^\ast \leftsample \encrypt(\pk, m^\ast)$
            \State $\hat{m} \leftsample \mathcal{A}^{\mathcal{O}_\texttt{ATK}}(
                1^\lambda, \pk, c^\ast
            )$
            \State \Return $\llbrack m^\ast = \hat{m} \rrbrack$
        \end{algorithmic}
    \end{algorithm}
}
\def\algpco{
    \begin{algorithm}[H]
        \caption*{$\pco(m \in \mathcal{M}, c \in \mathcal{C})$}
        \begin{algorithmic}[1]
            \State \Return $\llbrack \decrypt(\sk, c) = m \rrbrack$
        \end{algorithmic}
    \end{algorithm}
}
\begin{figure}[H]
    \centering
    \begin{minipage}[b]{0.49\textwidth}
        \algowatkgame
    \end{minipage}
    \begin{minipage}[b]{0.49\textwidth}
        \algpco
    \end{minipage}
    \caption{
        One-way security game of PKE (left) and plaintext-checking oracle (right)
    }\label{fig:ow-game-pco}
\end{figure}

In the \texttt{OW-ATK} game (see figure \ref{fig:ow-game-pco}), an adversary's goal is to recover the decryption of a randomly generated ciphertext. A challenger randomly samples a keypair and a challenge plaintext $m^\ast$, encrypts the challenge plaintext $c^\ast \leftsample \encrypt(\pk, m^\ast)$, then gives $\pk$ and $c^\ast$ to the adversary $A$. The adversary $A$, with access to some oracle $\mathcal{O}_\texttt{ATK}$, outputs a guess decryption $\hat{m}$. $A$ wins the game if its guess $\hat{m}$ is equal to the challenge plaintext $m^\ast$. The \textit{advantage} $\adv_\texttt{OW-ATK}$ of an adversary in this game is the probability that it wins the game:

@@

The capabilities of the oracle $\mathcal{O}_\texttt{ATK}$ depends on the choice of security goal $\texttt{ATK}$. Particularly relevant to our result is security against plaintext-checking attack (PCA), for which the adversary has access to a plaintext-checking oracle (PCO) (see figure \ref{fig:ow-game-pco}). A PCO takes as input a plaintext-ciphertext pair $(m, c)$ and returns \texttt{True} if $m$ is the decryption of $c$ or \texttt{False} otherwise.

@@

\begin{figure}[H]
    \centering
    \begin{minipage}[b]{0.49\textwidth}
        \begin{algorithm}[H]
            \caption*{KEM-IND-CCA2 game}
            \begin{algorithmic}[1]
                \State $(\pk, \sk) \leftsample \keygen(1^\lambda)$
                \State $(c^\ast, K_0) \leftsample \encap(\pk)$
                \State $K_1 \leftsample \mathcal{K}$
                \State $b \leftsample \{0, 1\}$
                \State $\hat{b} \leftsample A^{\mathcal{O}_\decap}(
                    1^\lambda, \pk, c^\ast, K_b
                )$
                \State \Return $\llbrack \hat{b} = b \rrbrack$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}\hfill
    \begin{minipage}[b]{0.49\textwidth}
        \begin{algorithm}[H]
        \caption*{$\mathcal{O}_\decap(c)$}
        \begin{algorithmic}[1]
            \State \Return $\decap(\sk, c)$
        \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \caption{IND-CCA2 game for KEM (left) and decapsulation oracle (right)}\label{fig:kem-game}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \begin{minipage}{0.7\textwidth}
        \begin{algorithm}[H]
            \caption*{\texttt{EUF-CMA} game}

            \begin{algorithmic}[1]
                \State $k^\ast \leftsample \mathcal{K}$
                \State $(\hat{m}, \hat{t}) \leftsample \mathcal{A}^{\mathcal{O}_\sign}()$
                \State \Return $
                    \llbrack \verify(k^\ast, \hat{m}, \hat{t}) 
                    \land (\hat{m}, \hat{t}) \not\in \mathcal{O}_\sign
                    \rrbrack
                $
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \caption{The existential forgery game}\label{fig:euf-cma-game}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \begin{minipage}[t]{0.49\textwidth}
        \begin{algorithm}[H]
            \caption*{$\kem\texttt{.}\keygen()$}
            \begin{algorithmic}[1]
                \State $(\pk, \sk^\prime) \leftsample \pke\texttt{.}\keygen()$
                \State $z \leftsample \mathcal{M}$
                \State $\sk \leftarrow (\sk^\prime, z)$
                    \Comment{$U^{\not\bot}, U^{\not\bot}_m$}
                \State $\sk \leftarrow \sk^\prime$
                    \Comment{$U^{\bot}, U^{\bot}_m$}
                \State \Return $(\pk, \sk)$
            \end{algorithmic}
        \end{algorithm}
        \begin{algorithm}[H]
            \caption*{$\kem\texttt{.}\encap(\pk)$}
            \begin{algorithmic}[1]
                \State $m \leftsample \mathcal{M}$
                \State $r \leftarrow G(m)$
                \State $c \leftarrow \pke\texttt{.}\encrypt(\pk, m, r)$
                \State $K \leftarrow H(m, c)$
                    \Comment{$U^\bot, U^{\not\bot}$}
                \State $K \leftarrow H(m)$
                    \Comment{$U_m^\bot, U_m^{\not\bot}$}
                \State \Return $(c, K)$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \begin{minipage}[t]{0.5\textwidth}
        \begin{algorithm}[H]
            \caption*{$\kem\texttt{.}\decap(\sk, c)$}
            \begin{algorithmic}[1]
                \Ensure $\sk = (\sk^\prime, z)$
                \Ensure $\sk^\prime$ is valid PKE secret key
                \Ensure $z$ is valid PKE plaintext
                \State $\hat{m} \leftarrow \pke\texttt{.}\decrypt(\sk^\prime, c)$
                \State $\hat{r} \leftarrow G(\hat{m})$
                \State $\hat{c} \leftarrow \pke\texttt{.}\encrypt(\pk, \hat{m}, \hat{r})$
                \If{$\hat{c} = c$}
                    \State $K \leftarrow H(\hat{m})$
                        \Comment{$U_m^\bot, U_m^{\not\bot}$}
                    \State $K \leftarrow H(\hat{m}, c)$
                        \Comment{$U^\bot, U^{\not\bot}$}
                \Else
                    \State $K \leftarrow H(z, c)$
                        \Comment{$U^{\not\bot}, U^{\not\bot}_m$}
                    \State $K \leftarrow \bot$
                        \Comment{$U^{\bot}, U^{\bot}_m$}
                \EndIf 
                \State \Return $K$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \caption{Summary of the modular Fujisaki-Okamoto transformation variations}\label{fig:fo-routines}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \begin{minipage}[t]{0.5\textwidth}
        \begin{algorithm}[H]
            \caption*{$\kem_\etm.\keygen()$}
            \begin{algorithmic}[1]
                \State $(\pk, \sk^\prime) \leftsample \pke\texttt{.}\keygen()$
                \State $z \leftsample \mathcal{M}$
                \State $\sk \leftarrow \sk^\prime \Vert z$
                \State \Return $(\pk, \sk)$
            \end{algorithmic}
        \end{algorithm}
        \begin{algorithm}[H]
            \caption*{$\kem_\etm.\encap(\pk)$}
            \begin{algorithmic}[1]
                \Ensure $\pk$ is some $\pke$ public key
                \State $m \leftsample \mathcal{M}$
                \State $k \leftarrow G(m)$
                \State $c^\prime \leftsample \pke\texttt{.}\encrypt(\pk, m)$
                \State $t \leftarrow \mac(k, c^\prime)$
                \State $K \leftarrow H(m, c^\prime)$
                \State $c \leftarrow c^\prime \Vert t$
                \State \Return $(c, K)$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}\hfill
    \begin{minipage}[t]{0.49\textwidth}
        \begin{algorithm}[H]
            \caption*{$\kem_\etm.\decap(\sk, c)$}
            \begin{algorithmic}[1]
                \Require $c = c^\prime \Vert t, \sk = sk^\prime \Vert z$
                \Ensure $c^\prime$ is some $\pke$ ciphertext
                \Ensure $t$ is some $\mac$ tag
                \Ensure $\sk^\prime$ is some $\pke$ secret key
                \Ensure $z$ is some $\pke$ plaintext
                \State $(c^\prime, t) \leftarrow c$
                \State $(\sk^\prime, z) \leftarrow \sk$
                \State $\hat{m} \leftarrow \pke\texttt{.}\decrypt(\sk^\prime, c^\prime)$
                \State $\hat{k} \leftarrow G(\hat{m})$
                \If{$\mac(\hat{k}, c^\prime) \neq t$}
                    \State $K \leftarrow H(z, c^\prime)$
                \Else
                    \State $K \leftarrow H(\hat{m}, c^\prime)$
                \EndIf
                \State \Return $K$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \caption{$\kem_\etm$ routines}\label{fig:etm-routines} 
\end{figure}

@@

\begin{figure}[H]
    \centering
    \begin{minipage}{0.6\textwidth}
    \begin{algorithm}[H]
        \caption*{$\pco(m, c)$}
        \begin{algorithmic}[1]
            \State $k \leftarrow G(m)$
            \State $t \leftarrow \mac(k, c)$
            \State \Return $\llbrack \mathcal{O}^\decap((c, t)) = H(m, c)\rrbrack$
        \end{algorithmic}
    \end{algorithm}
    \end{minipage}
    \caption{Every decapsulation oracle can be converted into a plaintext-checking oracle}\label{fig:pco-using-decap-oracle}
\end{figure}

@@

\subsection{Proof of theorem \ref{thm:ow-pca-implies-kem-ind-cca2}}
% TODO: write some intuition first?
\begin{proof}
    We will prove theorem \ref{thm:ow-pca-implies-kem-ind-cca2} using a sequence of games. The sequence of games is summarized in figure \ref{fig:etm-ind-cc2-sequence-of-games}.

    \begin{figure}[h]
        \centering
        \begin{minipage}[t]{0.5\textwidth}
            \begin{algorithm}[H]
                \caption*{\texttt{IND-CCA2} game for $\kem_\etm$}
                \begin{algorithmic}[1]
                    \State $(\pk, \sk) \leftsample \kem_\etm\texttt{.}\keygen()$
                    \State $m^\ast \leftsample \mathcal{M}$
                    \State $c^\prime \leftsample \pke\texttt{.}\encrypt(\pk, m^\ast)$
                    \State $k^\ast \leftarrow G(m^\ast)$
                        \Comment{Game 0-1}
                    \State $k^\ast \leftsample \mathcal{K}_\mac$
                        \Comment{Game 2-3}
                    \State $t \leftarrow \mac(k^\ast, c^\prime)$
                    \State $c^\ast \leftarrow c^\prime \Vert t$
                    \State $K_0 \leftarrow H(m^\ast, c^\prime)$
                        \Comment{Game 0-2}
                    \State $K_0 \leftsample \mathcal{K}_\kem$
                        \Comment{Game 3}
                    \State $K_1 \leftsample \mathcal{K}_\kem$
                    \State $b \leftsample \{0,1\}$
                    \State $\hat{b} \leftarrow A^{\mathcal{O}^\decap}(\pk, c^\ast, K_b)$
                        \Comment{Game 0}
                    \State $\hat{b} \leftarrow A^{\mathcal{O}^\decap_1}(\pk, c^\ast, K_b)$
                        \Comment{Game 1-3}
                    \State \Return $\llbrack \hat{b} = b \rrbrack$
                \end{algorithmic}
            \end{algorithm}
            \begin{algorithm}[H]
                \caption*{$\mathcal{O}^G(m)$}
                \begin{algorithmic}[1]
                    \If{$\exists (\tilde{m}, \tilde{k}) \in \mathcal{L}^G : \tilde{m} = m$}
                        \State \Return $\tilde{k}$
                    \EndIf
                    \State $k \leftsample \mathcal{K}_\mac$
                    \State $\mathcal{L}^G \leftarrow \mathcal{L}^G \cup \{(m, k)\}$
                    \State \Return $k$
                \end{algorithmic}
            \end{algorithm}
        \end{minipage}
        \begin{minipage}[t]{0.49\textwidth}
            \begin{algorithm}[H]
                \caption*{$\mathcal{O}^\decap(c)$}
                \begin{algorithmic}[1]
                    \State $(c^\prime, t) \leftarrow c$
                    \State $\hat{m} = \decrypt(\sk^\prime, c^\prime)$
                    \State $\hat{k} \leftarrow G(\hat{m})$
                    \If{$\mac(\hat{k}, c^\prime) = t$}
                        \State $K \leftarrow H(\hat{m}, c^\prime)$
                    \Else 
                        \State $K \leftarrow H(z, c^\prime)$
                    \EndIf 
                    \State \Return $K$
                \end{algorithmic}
            \end{algorithm}
            \vspace{-0.5cm}
            \begin{algorithm}[H]
                \caption*{$\mathcal{O}^\decap_1(c)$}
                \begin{algorithmic}[1]
                    \State $(c^\prime, t) \leftarrow c$
                    \If{$\exists (\tilde{m}, \tilde{k}) \in \mathcal{L}^G : 
                        \tilde{m} = \decrypt(\sk^\prime, c^\prime)
                        \land \mac(\tilde{k}, c^\prime) = t
                    $}
                        \State $K \leftarrow H(\tilde{m}, c^\prime)$
                    \Else
                        \State $K \leftarrow H(z, c^\prime)$
                    \EndIf
                    \State \Return $K$
                \end{algorithmic}
            \end{algorithm}
            \vspace{-0.5cm}
            \begin{algorithm}[H]
                \caption*{$\mathcal{O}^H(m, c)$}
                \begin{algorithmic}[1]
                    \If{$\exists (\tilde{m}, \tilde{c}, \tilde{K}) \in \mathcal{L}^H : \tilde{m} = m \land \tilde{c} = c$}
                        \State \Return $\tilde{K}$
                    \EndIf
                    \State $K \leftsample \mathcal{K}_\kem$
                    \State $\mathcal{L}^H \leftarrow \mathcal{L}^H \cup \{(m, c, K)\}$
                    \State \Return $K$
                \end{algorithmic}
            \end{algorithm}
        \end{minipage}
        \caption{Sequence of games}\label{fig:etm-ind-cc2-sequence-of-games}
    \end{figure}

@@

@@

A detailed description of $\texttt{K-PKE}$'s routines can be found in FIPS 203 (TODO: citation). The ``encrypt-then-MAC'' routines are described in figure \ref{fig:ml-kem-plus-routines}.

\begin{figure}[H]
    \centering

    \begin{minipage}[t]{0.5\textwidth}
        \begin{algorithm}[H]
            \caption*{$\texttt{ML-KEM}^\texttt{+}\texttt{.}\keygen\texttt{()}$}
            \begin{algorithmic}[1]
                \State $z \leftsample \{0,1\}^{256}$
                \State $(\pk, \sk^\prime) \leftsample \texttt{K-PKE}.\keygen()$
                \State $h \leftarrow H(\pk)$
                \State $\sk \leftarrow (sk^\prime \Vert \pk \Vert h \Vert z)$
                \State \Return $(\pk, \sk)$
            \end{algorithmic}
        \end{algorithm}\vspace{-0.5cm}
        \begin{algorithm}[H]
            \caption*{$\texttt{ML-KEM}^\texttt{+}\texttt{.}\encap\texttt{(\pk)}$}
            \begin{algorithmic}[1]
                \Require Public key $\pk$
                \State $m \leftsample \{0,1\}^{256}$
                \State $(\overline{K}, r, k) \leftarrow \texttt{XOF}(m \Vert H(\pk))$
                \State $c^\prime \leftarrow \texttt{K-PKE}.\encrypt(\pk, m, r)$
                \State $t \leftarrow \mac(k, c^\prime)$
                \State $K \leftarrow \texttt{KDF}(\overline{K} \Vert t)$
                \State $c \leftarrow (c^\prime, c)$
                \State \Return $(c, K)$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}\hfill
    \begin{minipage}[t]{0.49\textwidth}
        \begin{algorithm}[H]
            \caption*{$\texttt{ML-KEM}^\texttt{+}\texttt{.}\decap\texttt{(\sk, c)}$}
            \begin{algorithmic}[1]
                \Require Secret key $\sk = (\sk^\prime \Vert \pk \Vert h \Vert z)$
                \Require Ciphertext $c = (c^\prime \Vert t)$
                \State $(\sk^\prime, \pk, h, z) \leftarrow \sk$
                \State $(c^\prime, t) \leftarrow c$
                \State $\hat{m} \leftarrow \texttt{K-PKE}.\decrypt(\sk^\prime, c^\prime)$
                \State $(\overline{K}, \hat{r}, \hat{k}) \leftarrow \texttt{XOF}(\hat{m} \Vert h)$
                \State $\hat{t} \leftarrow \mac(\hat{k}, c^\prime)$
                \If{$\hat{t} = t$}
                    \State $K \leftarrow \texttt{KDF}(\overline{K} \Vert t)$
                \Else
                    \State $K \leftarrow \texttt{KDF}(z \Vert t)$
                \EndIf
                \State \Return $K$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}

    \caption{$\texttt{ML-KEM}^+$ routines}\label{fig:ml-kem-plus-routines}
\end{figure}
